Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2023-07-30 ba_resubmit_template_files/kundinger_resubmit3.tex
correct complication issue with pi at line 23 and 24; add comma after thus.
update references.
fix authors.

\documentclass[ba]{imsart}
%
\pubyear{2021}
\volume{TBA}
\issue{TBA}
% \doi{0000}
%\arxiv{}
\firstpage{1}
\lastpage{1}

@@

\begin{aug}
\author{\fnms{Brian} \snm{Kundinger}\thanksref{addr1}\ead[label=e1]{brian.kundinger@duke.edu}},
\author{\fnms{Jerome P.} \snm{Reiter}\thanksref{addr1}\ead[label=e2]{jreiter@duke.edu}}
\and
\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr2}\ead[label=e3]{beka@stat.duke.edu}}

@@

\address[addr1]{Department of Statistical Science,
  Duke University,
  P.O.\ Box 90251,
  Durham, NC 27708, USA
  \printead{e1}, % print email address of "e1"
  \printead*{e2}
}

\address[addr2]{Departments of Statistical Science and Computer Science,
  Duke University,
  P.O.\ Box 90251,
  Durham, NC 27708, USA
  \printead{e3}
}

@@

\begin{abstract}
%Recently, researchers have developed Bayesian versions of the Fellegi-Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational difficulties, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} can have markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
%
%\textcolor{blue}{We propose a variation on Bayesian Fellegi-Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.  Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}

\end{abstract}

@@

% Possible citation for bipartite matching. 
%\citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_A995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular for its mathematical simplicity. However, in many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. Analysts typically use an additional post-processing step to turn the list of linked records into a bipartite matching \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

@@

%Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review.

@@

%

%{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}}

@@

%Consider two data files $A$ and $B$, consisting of records $\{A_i: i = 1, \ldots, n_A\}$ and $\{B_j: j = 1, \ldots, n_B\}$. Suppose the files have $F$ fields in common to be used for linkage, and that these fields take on values $A_{if}$ and $B_{jf}$ respectively.

Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$. We define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$ as the collection of all record pairs $\gamma_{ij}$. 
%Additionally, we let $\Gamma_{ij}$ represent a random variable for the comparison vector for $A_i$ and $B_j$, such that $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.

@@

Instead of the prior over $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast beta linkage (\texttt{fabl}) prior below. For each $Z_j$, we have
%\begin{subequations}
\begin{align}
	p(Z_j = q| \pi) &= \begin{cases} 
	\frac{1}{n_A}\pi,  & q \leq n_A;\\
	1-\pi, &  q  = n_A + j;  \\
\end{cases} \label{eqn:fast_beta_prior}  \\
\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}). \nonumber
%\label{eqn:fast_beta_prior2}.
\end{align}
%\end{subequations}
We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match. The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $B$ that have matches in $A.$

@@

%\begin{table}[t!]
%	\centering
%	\begin{tabular}[t!]{ll}
%		Symbol & Description \\
%		\hline
%		$A, B$ & data files\\
%		$i \in 1, \ldots, n_A $ & index over records in $A$\\
%		$j \in 1, \ldots, n_B $ & index over records in $B$\\
%		$f \in 1, \ldots F$ & index over fields used for comparisons \\
%		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
%		$n_{AB}$ & number of matches between $A$ and $B$\\
%		$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
%		$Z_j = i$ & records $A_i$ and $B_j$ match \\
%		$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
%		$m_{fl}$ & $p(\gamma_{ij}^f = l | Z_j = i)$ \\
%		$u_{fl}$ & $p(\gamma_{ij}^f = l | Z_j \neq i)$ \\
%		$\pi$ & probability that a record in $B$ has a match in $A$ \\
%		\hline
%	\end{tabular}\caption{Summary of model notation.}\label{table_notation_A}
%\end{table}
%
For clarity, we present the full model for \texttt{fabl} below:
\begin{subequations}
\begin{align}
	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \gamma) &= \prod_{i=1}^{n_A}  \prod_{j=1}^{n_B}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}), \forall f = 1, \ldots, F, \label{eqn:m} \\
	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}),\forall f = 1, \ldots, F,  \label{eqn:u}\\
	p(Z_j = q| \pi)  &=
	\begin{cases} 
		\frac{1}{n_A}\pi,  & q \leq n_A; \\
		1-\pi, &  q  = n_A + j; \\
	\end{cases} \label{eqn:z}\\
	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi}.
\end{align}
\end{subequations}

%\hypertarget{posterior-sampling}{%
%	\subsection{Gibbs Sampler}
%	\label{gibbs_sampling}}

We estimate the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. Supplement \ref{app:derivations} presents derivations for the full conditional distributions. We initialize $\bm{m}$, $\bm{u}$ and $\pi$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$. Denote the current draw of the sampler by $(s)$ and the updated draw by $(s+1)$. To update $\bm{m}_f$ and $\bm{u}_f$ for all $f = 1, \ldots, F$, we sample
\begin{subequations}
\begin{align}
	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{u}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
	%where $\alpha_{fl}(\bm{Z}^{(s)})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i)$, and $\beta_{fl}(\bm{Z}^{(s)})= \\ \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i)$.
\end{align}
\end{subequations}
Next, we sample $\pi$ from its full conditional, given by
\begin{align}
	\pi^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)} \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
\end{align}
where $n_{AB}(\bm{Z}^{(s)}) = \sum_{j=1}^{n_B} I(Z_j^{(s)} \leq n_A)$.

Lastly, we sample $\bm{Z}$ componentwise from the full conditional for each $Z_j$:
\begin{align}
	\label{eqn:z_full_conditional}
	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)},  \pi^{(s+1)}\right) \propto
	\begin{cases} 
		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
		1 - \pi^{(s+1)}, & q  = n_A + j, \\
	\end{cases}
\end{align}
%\begin{align}
%\label{eqn:z_full_conditional}
%p\left(Z_j^{(s+1)}  = z_j|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
%\begin{cases} 
%	w_{z_j, j}^{(s+1)},  & z_j \leq n_A; \\
%	n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, & z_j  = n_A + j, \\
%\end{cases}
%\end{align}
where, for all $i \in \{1, \ldots, n_A\}$ and $j \in \{1, \ldots, n_B\}$, 
\begin{align}
	\label{eqn:fs_weight}
	w_{ij}^{(s)} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}^{(s)}}{u_{fl}^{(s)}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
\end{align}

@@

With \texttt{fabl}, we introduce two techniques to further expand the scalability of probabilistic record linkage. First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of storing the comparison vectors. Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches. 
%For convenience, Table \ref{table_notation_B} summarizes the notation introduced throughout this section.

%\begin{table}[t!]
%	\centering
%	\begin{tabular}[t!]{ll}
%		Symbol & Description \\
%		\hline
%		$P$ & number of agreement patterns exhibited in $\gamma$ \\
%		$h_p$ & agreement pattern $p$ \\
%		$e(h_p)$ & one-hot encoding of agreement pattern $p$ \\
%		$\gamma_{ij} = h_p$ & comparison vector between records $A_i$ and $B_j$ exhibits pattern $p$ \\
%		$r_{p_j}$ & list of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
%		$N_{p_j}$ & number of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
%		$N_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
%		$N_p$ & number of total comparison vectors that exhibit pattern $p$ \\
%
%		\hline
%	\end{tabular}\caption{Summary of hashing notation.}\label{table_notation_B}
%\end{table}

@@

We then map the integers in \eqref{eqn:hashing} to sequential integers $\{1, \ldots, P\}$. Denote each unique agreement pattern as $h_p = (h_p^1, \ldots, h_p^F)$, and the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$. When the $(i,j)$ record pair exhibits agreement pattern $p$, we say $\gamma_{ij} = h_p$. In calculations, we will at times use the one-hot encoding of agreement pattern $h_p$, denoted $e(h_p)$. This is a vector of length $\sum_{f=1}^F L_f$ in which the $l + \sum_{k=1}^{f-1} L_k$ component is 1 when $\gamma_{ij}^f = l$, and 0 otherwise. 
%A more complete example of the one-hot encoding is provided in Supplement \ref{app:ohe}.

@@

%We also use (\ref{likelihood_efficient}) to sample $\bm{Z}$ efficiently.

@@

We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
\begin{align}
	\label{eqn:gibbs1}
	p\left( \Gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
	\begin{cases} 
		%\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & \gamma_{Z_j^{(s+1)}, j} = h_p; \\
		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & h_p \in \mathcal{P}; \\
		1- \pi^{(s+1)} , &   \text{otherwise}. \\
	\end{cases}
\end{align}
Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
\begin{align}
	\label{eqn:gibbs2}
	p\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
		\frac{1}{N_{p_j}}, & q \in r_{p_j}; \\
		0, & \text{otherwise.} \\
	\end{cases}
\end{align}

@@

		 %For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records.

@@

%Let $n_a$ and $n_b$ be the number of records in batches $A^a$ and $B^b$ respectively. Instead of storing $n_A n_B$ record labels, with SEI we store at most $t_A \sum_{b = 1}^{t_B} n_b  P  S$ labels.

@@

%	\begin{figure}[t]
%		\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_speed_plot_new} 
%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$, as described in Section \ref{speed}. We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
%		\end{center}
%	\end{figure}
%
%	\begin{figure}[h!]
%	\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/speed_plot_fixed_nB_new} 
%		\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_B$ fixed at 500, as described in Section \ref{speed}. We see near linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
%	\end{center}
%\end{figure}

@@

%		\begin{figure}
%		\begin{minipage}[c]{0.45\linewidth}
%			\begin{center}
%				\includegraphics[width=0.8\textwidth]{../notes/figures/sadinle_speed_plot_new}
%				\caption{Simulations with increasing values of both $n_A$ and $n_B$. We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
%			\end{center}
%		\end{minipage}
%		\hfill
%		\begin{minipage}[c]{0.45\linewidth}
%			\begin{center}
%				\includegraphics[width=0.8\textwidth]{../notes/figures/speed_plot_fixed_nB_new} 
%				\caption{Simulations with $n_B$ = 500, We see near linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
%			\end{center}
%		\end{minipage}%
%	\end{figure}

@@

	We compare \texttt{fabl} to \texttt{BRL} in terms of recall, precision and F-measure, as defined in \cite{christen_2012}. Recall is the proportion of true matches found by the model, that is, $\sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j, Z_j \leq n_A) / \sum_{j=1}^{n_B} I(Z_j \leq n_A)$. Precision is the proportion of links found by the model that are true matches, that is, $\sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j, Z_j \leq n_A) / \sum_{j=1}^{n_B} I(\hat{Z}_j \leq n_A)$. The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2  (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$. In Figure \ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap. 
	%In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse than \texttt{BRL} on average; however, the overall accuracy level remains high.

@@

	%It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.

@@

	\begin{table}[!t]
		\begin{tabular}[t]{llllll}
			\multicolumn{2}{c}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
			\cline{3-6}
			Fields & Similarity & 1 & 2 & 3 & 4\\
			\hline
			First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, .5] & (.5, 1]\\
			Year of Death & Absolute Difference & 0 & 1 & 2 & 3+\\
			Month of Death & Absolute Difference & 0 & 1 & 2-3 & 4+\\
			Day of Death & Absolute Difference & 0 & 1-2 & 3-7 & 8+\\
			Municipality & Binary & Agree & Disagree &  & \\
			\hline
		\end{tabular}
		\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \cite{sadinle_bayesian_2017}. This setup leads to 1875 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}
		%	\end{table}

>>>>

\documentclass[ba]{imsart}
%
\pubyear{2023}
\volume{TBA}
\issue{TBA}
% \doi{0000}
%\arxiv{}
\firstpage{1}
\lastpage{1}

@@

\begin{aug}
\author{\fnms{Brian} \snm{Kundinger}\thanksref{addr1}\ead[label=e1]{brian.kundinger@duke.edu}},
\author{\fnms{Jerome P.} \snm{Reiter}\thanksref{addr1}\ead[label=e2]{jreiter@duke.edu}}
\and
\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr1}\ead[label=e3]{beka@stat.duke.edu}}

@@

\address[addr1]{Department of Statistical Science,
  Duke University,
  P.O.\ Box 90251,
  Durham, NC 27708, USA
  \printead{e1}, % print email address of "e1"
  \printead*{e2},
   \printead*{e3}
}

%\address[addr2]{Departments of Statistical Science and Computer Science,
%  Duke University,
%  P.O.\ Box 90251,
%  Durham, NC 27708, USA
%  \printead{e3}
%}

@@

\begin{abstract}
Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational difficulties, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} can have markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
\end{abstract}

@@

\newpage
The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular for its mathematical simplicity. However, in many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. Analysts typically use an additional post-processing step to turn the list of linked records into a bipartite matching \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

@@

@@

@@

Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$. We define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$ as the collection of all record pairs $\gamma_{ij}$.

@@

Instead of the prior over $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast beta linkage (\texttt{fabl}) prior below. For each $Z_j$, we have

\begin{align}
	p(Z_j = q| \pi) &= \begin{cases} 
	\frac{1}{n_A}\pi,  & q \leq n_A;\\
	1-\pi, &  q  = n_A + j;  \\
\end{cases} \label{eqn:fast_beta_prior}  \\
\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}). \nonumber
\end{align}

We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match. The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $B$ that have matches in $A.$

@@

For clarity, we present the full model for \texttt{fabl} below:
\begin{subequations}
\begin{align}
	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \gamma) &= \prod_{i=1}^{n_A}  \prod_{j=1}^{n_B}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}), \forall f = 1, \ldots, F, \label{eqn:m} \\
	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}),\forall f = 1, \ldots, F,  \label{eqn:u}\\
	p(Z_j = q| \pi)  &=
	\begin{cases} 
		\frac{1}{n_A}\pi,  & q \leq n_A; \\
		1-\pi, &  q  = n_A + j; \\
	\end{cases} \label{eqn:z}\\
	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi}.
\end{align}
\end{subequations}

We estimate the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. Supplement \ref{app:derivations} presents derivations for the full conditional distributions. We initialize $\bm{m}$, $\bm{u}$ and $\pi$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$. Denote the current draw of the sampler by $(s)$ and the updated draw by $(s+1)$. To update $\bm{m}_f$ and $\bm{u}_f$ for all $f = 1, \ldots, F$, we sample
\begin{subequations}
\begin{align}
	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{u}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
\end{align}
\end{subequations}
Next, we sample $\pi$ from its full conditional, given by
\begin{align}
	\pi^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)} \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
\end{align}
where $n_{AB}(\bm{Z}^{(s)}) = \sum_{j=1}^{n_B} I(Z_j^{(s)} \leq n_A)$.

Lastly, we sample $\bm{Z}$ componentwise from the full conditional for each $Z_j$:
\begin{align}
	\label{eqn:z_full_conditional}
	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)},  \pi^{(s+1)}\right) \propto
	\begin{cases} 
		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
		1 - \pi^{(s+1)}, & q  = n_A + j, \\
	\end{cases}
\end{align}
where, for all $i \in \{1, \ldots, n_A\}$ and $j \in \{1, \ldots, n_B\}$, 
\begin{align}
	\label{eqn:fs_weight}
	w_{ij}^{(s)} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}^{(s)}}{u_{fl}^{(s)}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
\end{align}

@@

With \texttt{fabl}, we introduce two techniques to further expand the scalability of probabilistic record linkage. First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of storing the comparison vectors. Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.

@@

We then map the integers in \eqref{eqn:hashing} to sequential integers $\{1, \ldots, P\}$. Denote each unique agreement pattern as $h_p = (h_p^1, \ldots, h_p^F)$, and the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$. When the $(i,j)$ record pair exhibits agreement pattern $p$, we say $\gamma_{ij} = h_p$. In calculations, we will at times use the one-hot encoding of agreement pattern $h_p$, denoted $e(h_p)$. This is a vector of length $\sum_{f=1}^F L_f$ in which the $l + \sum_{k=1}^{f-1} L_k$ component is 1 when $\gamma_{ij}^f = l$, and 0 otherwise.

@@

@@

We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
\begin{align}
	\label{eqn:gibbs1}
	p\left( \Gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
	\begin{cases} 
		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & h_p \in \mathcal{P}; \\
		1- \pi^{(s+1)} , &   \text{otherwise}. \\
	\end{cases}
\end{align}
Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
\begin{align}
	\label{eqn:gibbs2}
	p\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
		\frac{1}{N_{p_j}}, & q \in r_{p_j}; \\
		0, & \text{otherwise.} \\
	\end{cases}
\end{align}

@@

@@

@@

@@

@@

	We compare \texttt{fabl} to \texttt{BRL} in terms of recall, precision and F-measure, as defined in \cite{christen_2012}. Recall is the proportion of true matches found by the model, that is, $\sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j, Z_j \leq n_A) / \sum_{j=1}^{n_B} I(Z_j \leq n_A)$. Precision is the proportion of links found by the model that are true matches, that is, $\sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j, Z_j \leq n_A) / \sum_{j=1}^{n_B} I(\hat{Z}_j \leq n_A)$. The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2  (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$. In Figure \ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.

@@

@@

	\begin{table}[!t]
		\begin{tabular}[t]{llllll}
			\multicolumn{2}{c}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
			\cline{3-6}
			Fields & Similarity & 1 & 2 & 3 & 4\\
			\hline
			First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, .5] & (.5, 1]\\
			Year of Death & Absolute Difference & 0 & 1 & 2 & 3+\\
			Month of Death & Absolute Difference & 0 & 1 & 2-3 & 4+\\
			Day of Death & Absolute Difference & 0 & 1-2 & 3-7 & 8+\\
			Municipality & Binary & Agree & Disagree &  & \\
			\hline
		\end{tabular}
		\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \cite{sadinle_bayesian_2017}. This setup leads to 1875 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}

		\centering
		\begin{tabular}[t]{lllll}
			\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
			\cline{3-5}
			Fields & Similarity & 1 & 2 & 3\\
			\hline
			First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, 1]\\
			Year of Death & Binary & Agree & Disagree & \\
			Month of Death & Binary & Agree & Disagree & \\
			Day of Death & Absolute Difference & 0 & 1 & 2+\\
			Municipality & Binary & Agree & Disagree & \\
			\hline
		\end{tabular}
		\caption{Construction of comparison vectors for El Salvador data for increased speed under \texttt{fabl}. This setup leads to 432 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_2}
	\end{table}

	\begin{figure}[p]
	\begin{center}
		\includegraphics[width=0.99\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
		\caption{Posterior distribution and Bayes estimate of overlap across the two files in the El Salvador case study. We note they are quite similar under both methods.}
		\label{fig:overlap-plot}
	\end{center}

	\begin{center}
		\includegraphics[width=0.99\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
		\caption{Posterior estimates of $\bm{m}$ parameters with 95\% credible intervals for the El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
		\label{fig:m-and-u}
	\end{center}
\end{figure}
